\chapter*{Acknowledgements}
First and foremost, I would like to thank my supervisor Farid Ould-Saada for introducing me to particle physics and helped me find both an exciting and challenging project. My co-supervisor, Eirik Gramstad, have been a great support and help when errors have occurred during the work done for this thesis. Thank you both for never giving up on me and for dragging me over the finish line.

A big thank you goes to the HEP-group which have welcomed me with open arms since the day I started at HEP in 2018. I appreciate the great social environment and that all of you are always ready to answer my questions, both big and small. And of course, thank you for making me a quiz nerd.

Even though the whole HEP-group probably should have been named, I need to give an extra thank you to Eli, Helén and Oda! You have been with me on the highest highs and lowest lows, so thank you for always encouraging, motivating, pushing, and also feeding me. 

Of course a huge thank you goes to "my brother from another mother", Federico! You have been a huge support both on and off campus, and I am grateful for sharing this journey with you. Particle physics brought us together and started our friendship, and even though we are now going different ways, I hope it will continue forever. I would also like to thank both you and Jake for a lot of memorable times this last year, and pushing me to be more confident in another language.

Knut Oddvar Høie Vadla, thank you! Without your expertise and patience, I would never have finished this master thesis. You have always taken the time to explain things to me, even when at your new job. To know that you would always be willing to help, have been a relief in challenging situations. 

Last but not least, I would like to thank my family, especially my parents and my sister. You have always been nothing but proud of me, and having a sister close by during this degree have been an invaluable support. Thank you very much.

The research presented in this thesis has benefited from the Experimental Infrastructure for Exploration of Exascale Computing (eX3), which is financially supported by the Research Council of Norway under contract 270053.
