\chapter{The Standard Model and Beyond}
\label{sec:Theory}
Although the Standard Model (SM) of particle physics is one of the greatest triumphs of modern physics, it cannot describe all physical phenomena observed in nature. There are several shortcomings of the this model, which implies that we need an extension of the SM to understand and explain these problems. There are several different suggested theoretical solutions to these, and one of them is called supersymmetry (SUSY). SUSY introduces multiple new particles as an extension to the SM particles, but no SUSY particle has so far been detected. Many supersymmetric theories inlcude a weakly interacting massive particle (WIMPs) and thus a candidate to explain the dark matter (DM) in our Universe, a brief overview of which will be provided in this thesis. Before we discuss SUSY and DM, we begin with a run-through of the SM.
\input{Sections/Theory/SM.tex}


\section{Beyond the Standard Model}
As mentioned earlier, the SM is not enough to explain every phenomena observed in experiments. There are several challenging aspects of the SM; for instance, many of the parameters in the model are made to fit experimental data and do not come from theoretical principles. 
The SM does not offer a solution to unify the framework with gravity. The large difference between the weak energy scale and the Planck scale, known as the hierarchy problem, is another shortcoming of the SM. Moreover, the SM does not provide any explanation for the dark matter observed in our Universe nor does it explain the small, but non-zero, masses of neutrinos.  

Given these shortcomings, various theories have been proposed addressing some of the open questions. Several extensions of the Standard Model exist, and there are pros and cons to all of them. Furthermore, even though the search has been going on for years, no concluding scientific evidence has been found to favor one particular SM extension over another. Below we detail an outline of one possibility:
Supersymmetry, which is often denoted SUSY for short.

\input{Sections/Theory/SUSY.tex}

\input{Sections/Theory/DM.tex}